---
title: "Batch Re-name Files"
output: github_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(reprex)
knitr::opts_chunk$set(
    warning = FALSE,
    message = FALSE,
    fig.path = "img/",
    fig.width = 7.252,
    fig.height = 4,
    comment = "#>",
    fig.retina = 3 # Better figure resolution
)
```

## Motivation: Change *a lot* of file names

Assume I have the following files: 

```{r dir_tree}
fs::dir_tree("img")
```

I want to change the prefix (`00-`) to `01-` for all of the files. I can do this with `fs` and `purrr`

## File information

```{r dir_files}
dir_files <- fs::dir_info("img", regexp = "[.]png$") %>% 
    filter(type == "file") %>% 
    select(path, type, ends_with("time"))
glimpse(dir_files)
```

Create the base name:

```{r file_names}
file_names <- dir_files %>% 
    mutate(file_name = basename(path)) %>% 
    select(path, file_name, everything())
glimpse(file_names)
```

Now we replace all prefixes: 

```{r new_file_name}
new_file_names <- file_names %>% 
    mutate(new_file_name = 
               str_replace_all(file_name,
                pattern = "^00-", 
                replacement = "01-")) %>% 
    select(path, file_name, new_file_name, everything())
glimpse(new_file_names)
```

Add the original file path back to the new file names:

```{r new_paths}
new_paths <- new_file_names %>% 
    mutate(base_path = str_remove_all(string = path, pattern = file_name),
           new_path = paste0(base_path, new_file_name)) %>% 
    select(path, new_path, file_name, new_file_name, ends_with("time"))
glimpse(new_paths)
```


## Do it for one element 

Now we can extract one element and rename it to see if it works: 

```{r img-test}
fs::dir_create("img-test")
fs::file_copy(path = "raw-images/00-538-example-01.png", 
              new_path = "img-test/00-538-example-01.png", 
              overwrite = TRUE)
fs::dir_tree("img-test")
```

```{r test-1}
file.rename(from = "img-test/00-538-example-01.png", 
            to = "img-test/01-538-example-01.png")
```

We can check to verify 

```{r dir_tree-test-1}
fs::dir_tree("img-test", regexp = "01-")
```

```{r unlink, echo=FALSE}
unlink(x = "img-test/01-538-example-01.png", force = TRUE)
```

Now we do it for all the new names in `new_path`. 

```{r old_paths}
old_file_paths <- new_paths$path
head(old_file_paths)
```

```{r new_file_paths}
new_file_paths <- new_paths$new_path
head(new_file_paths)
```

```{r map2}
map2(.x = old_file_paths, .y = new_file_paths, file.rename)
```

Verify 

```{r dir_tree-verify}
fs::dir_tree("img", regexp = "01-")
```

## Reset `img/`

```{r replace-img}
# remove contents of img/
fs::dir_delete("img")
# get 00- files from raw-images/
img_00_files <- fs::dir_ls("raw-images/", regexp = "/00-")
# create img again
new_img_path <- fs::dir_create("img/")
# copy files over 
purrr::quietly(map2(.x = img_00_files, .y = new_img_path, 
     .f = fs::file_copy, overwrite = TRUE))
# verify
fs::dir_tree("img")
```

## `batch_rename_files()`

Bundle as a function: 

```{r batch_rename_files}
batch_rename_files <- function(path, pattern, replace, prefix, extension) {
    
    # get file_ext
    file_ext <- paste0("[.]", extension, "$")
    
    # get dir_files
    dir_files <- fs::dir_info(path = path, regexp = file_ext) 
    dir_files <- dplyr::filter(dir_files, type == "file") 
    dir_files <- dplyr::select(dir_files, path, type, ends_with("time"))
    
    # define pattern
    regex_pattern <- as.character(pattern)
    # define replace
    regex_replace <- as.character(replace)
    
        # prefix
    if (prefix == TRUE) {
        
        regex <- paste0("^", regex_pattern)
         # pattern
    } else { 
        
        regex <- as.character(regex_pattern)
       
    } 

    file_names <- mutate(dir_files, file_name = basename(path))
    
    new_file_names <- mutate(file_names, 
           new_file_name = 
                    stringr::str_replace_all(file_name,
                    pattern = regex, 
                    replacement = regex_replace))
    
    new_paths <- mutate(new_file_names, 
                    base_path = 
                    stringr::str_remove_all(string = path, pattern = file_name),
                    new_path = paste0(base_path, new_file_name)) %>% 
    select(path, new_path, file_name, new_file_name, ends_with("time"))
    
    old_file_paths <- new_paths$path
    new_file_paths <- new_paths$new_path
    
    # return(new_file_paths)
    
    purrr::quietly(purrr::map2(.x = old_file_paths, .y = new_file_paths,
                               file.rename))
    
}
```

## Set up tests 

```{r test_01_path, results='hide'}
test_01_path <- "test-01/"
fs::dir_create(test_01_path)
test_01_files <- fs::dir_ls("raw-images/", regexp = "/00-")
purrr::quietly(map2(test_01_files, .y = test_01_path, 
     .f = fs::file_copy, overwrite = TRUE))
```

```{r test_02_path, results='hide'}
test_02_path <- "test-02/"
fs::dir_create(test_02_path)
test_02_files <- fs::dir_ls("raw-images/", regexp = "-1[.]png$")
purrr::quietly(map2(test_02_files, .y = test_02_path, 
     .f = fs::file_copy, overwrite = TRUE))
```

## Test 1

Test renaming `00-` to `01-`:

```{r test-01-batch_rename_files}
batch_rename_files(path = "test-01", pattern = "00-", replace = "01-", prefix = TRUE, extension = "png")
```

```{r verify-test-01}
fs::dir_tree("test-01")
```

## Reset `test-01/`

```{r reset-test-01}
# remove contents of img/
fs::dir_delete("test-01/")
# get 00- files from raw-images/
img_00_files <- fs::dir_ls("raw-images/", regexp = "/00-")
# create img again
test_01_img_path <- fs::dir_create("test-01/")
# copy files over 
purrr::quietly(map2(.x = img_00_files, .y = test_01_img_path, 
     .f = fs::file_copy, overwrite = TRUE))
# verify
fs::dir_tree("test-01/")
```


## Test 2

Test renaming `-1.png` to `.png`:

```{r batch_rename_files-test-2}
batch_rename_files(path = "test-02", pattern = "-1", replace = "", prefix = FALSE, extension = "png")
```

```{r verify-test-02}
fs::dir_tree("test-02")
```

## Reset `test-02/`

```{r reset-test-02}
# remove contents of img/
fs::dir_delete("test-02/")
# get 00- files from raw-images/
img_1_suffix_files <- fs::dir_ls("raw-images/", regexp = "-1[.]png$")
# create img again
test_02_img_path <- fs::dir_create("test-02/")
# copy files over 
purrr::quietly(map2(.x = img_1_suffix_files, .y = test_02_img_path, 
     .f = fs::file_copy, overwrite = TRUE))
# verify
fs::dir_tree("test-02/")
```